\documentclass[11pt]{amsart}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}

\newtheorem{theorem}{Theorem}

\title{The Burkholder Martingale Transform Inequality in Lean}
\author{}
\date{}

\begin{document}

\maketitle

\section{Overview}

This note describes the formal Lean theorem
\[
  \texttt{Lp\_Burkholder\_inequality\_martingaleTransform}
\]
proved in \texttt{MartingaleTransforms.lean}.  The result is the real-valued
martingale transform inequality
\[
  \norm{(w \star_m f)_n}_{L^p(\mu)}
    \le (p^\ast - 1)\,\norm{f_n}_{L^p(\mu)}
\]
for every time \(n\), where \(1 < p < \infty\), \(f\) is a martingale, and
\(w\) is a predictable multiplier bounded by \(1\) in absolute value.

In Lean the theorem is stated with the extended nonnegative exponent
\(p : \mathbb{R}_{\ge 0,\infty}\).  The hypotheses include \(1 < p\),
\(p \ne \infty\), and the proof uses \(p.toReal\) for the real exponent in
the Burkholder functions.

\section{Basic Definitions}

\subsection{Martingale differences}

For a discrete real-valued process \(f : \N \to \Omega \to \R\), the
martingale difference sequence is defined by
\[
  d_0 f = f_0,\qquad
  d_{n+1} f = f_{n+1} - f_n.
\]
In Lean this is
\[
\texttt{martingaleDiff f 0 = f 0},
\qquad
\texttt{martingaleDiff f (n+1) = f (n+1) - f n}.
\]

\subsection{Martingale transforms}

Given a multiplier process \(v : \N \to \Omega \to \R\), the discrete
martingale transform of \(f\) by \(v\) is
\[
  (v \star_m f)_n
    = \sum_{i=0}^n v_i\, d_i f.
\]
The Lean definition is
\[
\texttt{martingaleTransform v f :=
  fun n => (Finset.range (n + 1)).sum fun i => v i * martingaleDiff f i}.
\]
The notation used in the file is
\[
  v \star_m f.
\]
The increment identity proved in Lean is
\[
  (v \star_m f)_{n+1} - (v \star_m f)_n
    = v_{n+1}(f_{n+1}-f_n).
\]

\subsection{Predictability}

The multiplier \(w\) is assumed to be strongly predictable with respect to a
filtration \(\F\):
\[
  \texttt{IsStronglyPredictable }\F\,w.
\]
This is the formal version of the usual condition that \(w_{n+1}\) is
\(\F_n\)-measurable.  Predictability is what allows the proof to pull the
multiplier out of conditional expectations when showing that the linear terms
have expectation zero.

\subsection{The auxiliary transforms \(X\) and \(Y\)}

The proof uses two transforms associated to \(w\):
\[
  w^+_n = w_n + 1,\qquad w^-_n = w_n - 1.
\]
In Lean these are \texttt{plusOne w} and \texttt{minusOne w}.  The notation
\[
  X_n[w,f] = ((w+1)\star_m f)_n,\qquad
  Y_n[w,f] = ((w-1)\star_m f)_n
\]
is introduced by scoped notation.  These satisfy the algebraic identities
\[
  \frac{X_n+Y_n}{2} = (w\star_m f)_n,
  \qquad
  \frac{X_n-Y_n}{2} = f_n.
\]
The final theorem uses these identities together with the definition of the
Burkholder function \(v\).

\section{The Main Martingale Transform Theorem}

The Lean theorem has the following mathematical content.

\begin{theorem}[Burkholder martingale transform inequality]
Let \((\Omega,\mathcal{A},\mu)\) be a finite measure space and let
\((\F_n)_{n\in\N}\) be a filtration.  Let \(1 < p < \infty\).  Suppose
\(f = (f_n)\) is a real-valued martingale, \(f_n \in L^p(\mu)\) for every
\(n\), and \(w = (w_n)\) is strongly predictable with
\[
  |w_n(\omega)| \le 1
\]
for every \(n\), almost everywhere.  Then for every \(n\),
\[
  \norm{(w\star_m f)_n}_{L^p(\mu)}
    \le \bigl(p^\ast-1\bigr)\norm{f_n}_{L^p(\mu)},
\]
where
\[
  q = \frac{p}{p-1},
  \qquad
  p^\ast = \max\{p,q\}.
\]
\end{theorem}

In Lean, the constant is written as
\[
  \texttt{ENNReal.ofReal (Burkholder.pStar p.toReal - 1)}.
\]

\section{The Burkholder Function \(v\)}

The proof is organized around the Burkholder expression
\[
  v(p,x,y)
    =
      \left|\frac{x+y}{2}\right|^p
      -
      |p^\ast-1|^p
      \left|\frac{x-y}{2}\right|^p.
\]
This is the Lean definition \texttt{Majorants.v}.  Substituting
\[
  x = X_n[w,f], \qquad y = Y_n[w,f]
\]
and using the identities above gives
\[
  v(p,X_n,Y_n)
    =
      |(w\star_m f)_n|^p
      -
      |p^\ast-1|^p |f_n|^p.
\]
Thus the inequality
\[
  \int v(p,X_n,Y_n)\,d\mu \le 0
\]
is exactly the \(p\)-power form of the martingale transform estimate.
The last step of the Lean proof unfolds the definition of \(v\), converts the
integral inequality into an inequality of real \(L^p\)-norms, and then converts
back to \texttt{eLpNorm}.

\section{The Majorant Theorem \texttt{exists\_majorant\_p\_g\_1}}

The analytic heart of the formalization is
\[
  \texttt{exists\_majorant\_p\_g\_1}.
\]
For every real \(p>1\), it produces functions
\[
  u,\quad u_x,\quad u_y : \R \to \R \to \R
\]
and a constant \(C\ge 0\) such that:

\begin{itemize}
  \item \(u,u_x,u_y\) are continuous on all of \(\R^2\);
  \item \(u\) has \(p\)-growth:
    \[
      |u(x,y)| \le C(|x|^p+|y|^p);
    \]
  \item the selected derivatives have \((p-1)\)-growth:
    \[
      |u_x(x,y)|,\ |u_y(x,y)|
        \le C(|x|^{p-1}+|y|^{p-1});
    \]
  \item \(u\) satisfies the tangent/concavity inequality in martingale-transform
    directions:
    \[
      hk\le 0
      \Longrightarrow
      u(x+h,y+k)
        \le u(x,y)+u_x(x,y)h+u_y(x,y)k;
    \]
  \item \(u\) majorizes \(v\):
    \[
      v(p,x,y)\le u(x,y);
    \]
  \item \(u\) is nonpositive on the region \(xy\le 0\):
    \[
      xy\le 0 \Longrightarrow u(x,y)\le 0.
    \]
\end{itemize}

\subsection{Why the majorant \(u\) is important}

The role of the majorant \(u\) is to turn a norm inequality into a
supermartingale-type inequality.  The function \(v\) directly contains the
desired estimate, but by itself it does not have the concavity property needed
to use the martingale structure.  The majorant \(u\) supplies this missing
property:

\begin{enumerate}
  \item Since \(v\le u\), it is enough to prove that the expectation of
    \(u(X_n,Y_n)\) is nonpositive.
  \item The tangent inequality for \(u\) shows that
    \[
      u(X_{n+1},Y_{n+1})
        \le u(X_n,Y_n)
          + u_x(X_n,Y_n)(X_{n+1}-X_n)
          + u_y(X_n,Y_n)(Y_{n+1}-Y_n).
    \]
  \item The condition \(|w|\le 1\) implies
    \[
      (X_{n+1}-X_n)(Y_{n+1}-Y_n)\le 0,
    \]
    which is exactly the hypothesis needed for the tangent inequality.
  \item By predictability of \(w\) and the martingale property of \(f\), the
    linear terms have integral zero.
  \item Therefore,
    \[
      \int u(X_{n+1},Y_{n+1})\,d\mu
        \le
      \int u(X_n,Y_n)\,d\mu.
    \]
  \item At the initial time, \(X_0Y_0\le 0\), so \(u(X_0,Y_0)\le 0\).
    Hence \(\int u(X_n,Y_n)\,d\mu\le 0\) for every \(n\).
\end{enumerate}

Combining this with \(v\le u\) gives
\[
  \int v(p,X_n,Y_n)\,d\mu \le 0,
\]
and the definition of \(v\) then yields the desired martingale transform
inequality.

\end{document}
